---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# aof package

[![Build Status](https://travis-ci.org/frareb/osfi.svg?branch=master)](https://travis-ci.org/frareb/osfi)
[![CRAN version](https://www.r-pkg.org/badges/version/aof)](https://CRAN.R-project.org/package=aof)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/aof)](https://CRAN.R-project.org/package=aof)
[![Coverage Status](https://img.shields.io/codecov/c/gh/frareb/osfi/master.svg)](https://codecov.io/gh/frareb/osfi?branch=master)

## Purpose of the package

A breakpoint-based method to detect ontogenetic shifts in univariate time-activity budget series of central-place foraging insects. The method finds a single breakpoint according to the likelihood function. The method was developed with honey bees in order to detect the Age at Onset of Foraging (AOF), but can be used for the detection of other ontogenetic shifts in other central-place foraging insects. 

## Installation instructions

```{r inst01, eval=FALSE}
# from CRAN
install.packages("aof")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("frareb/aof")
```

## Authors' affiliation

UMR EGCE (IRD, CNRS, Univ. Paris-Saclay), Gif-sur-Yvette, France

To cite this package please use: 

```{r auth01, eval=TRUE}
citation("aof") 
```
