---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# aof package

[![Build Status](https://travis-ci.org/frareb/aof.svg?branch=master)](https://travis-ci.org/frareb/aof)
[![CRAN version](https://www.r-pkg.org/badges/version/aof)](https://CRAN.R-project.org/package=aof)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/aof)](https://CRAN.R-project.org/package=aof)
[![Coverage Status](https://img.shields.io/codecov/c/gh/frareb/aof/master.svg)](https://codecov.io/gh/frareb/aof?branch=master)

## Purpose of the package

A breakpoint-based method to detect ontogenetic shifts in univariate time-activity budget series of central-place foraging insects. The method finds a single breakpoint according to the likelihood function. 

## Installation instructions

```{r inst01, eval=FALSE}
# from CRAN
install.packages("aof")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("frareb/aof")
```

## Authors' affiliation

UMR EGCE (IRD, CNRS, Univ. Paris-Saclay), Gif-sur-Yvette, France

To cite this package please use: 

```{r auth01, eval=TRUE}
citation("aof") 
```
